---
title: "Assignment 3"
author: "Husayn Jessa and Aidan Bodner"
date: "16/11/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages
```{r Packages}
library(here)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(ggrepel)
library(ggtext)
library(showtext)
library(stringr)
library(fastDummies)
library(modelsummary)
```

# Data and Cleaning
```{r Data and Cleaning}
name <- Sys.info() 
name[7]

assign_data <- read.csv("a3_p1_sourdough_trends.csv")

```

# Question 1a
```{r Question 1a}

assign_data$date <- str_replace_all(assign_data$date, "T00:00:00Z", "")
assign_data$date <- as.factor(assign_data$date)
assign_data$date <- as.numeric(assign_data$date) 

q1_line_graph <- ggplot(assign_data, (aes(x = date, y = hits, group = keyword))) + 
  geom_line(aes(color = keyword)) + theme_linedraw() + labs(x = "Date", y = "Hits") +
  geom_vline(xintercept = as.numeric(assign_data$date[75]), 
    color = "red") + geom_text(aes(x = as.numeric(date[120]), label="Start of Pandemic (2020-03-15)", y=90), color = "red") +
  scale_x_discrete(breaks = c("2020-01-01",
                              "2020-02-01",
                              "2020-03-01",
                              "2020-04-01",
                              "2020-05-01",
                              "2020-06-01",
                              "2020-07-01",
                              "2020-08-01"), labels = waiver())

q1_line_graph


#I wanted to make a line graph without converting our dates to numeric:
assign_data$date <- str_replace_all(assign_data$date, "T00:00:00Z", "")

hq1_line_graph <- ggplot(assign_data, aes(x = date, y = hits, colour = keyword, group = keyword)) +
                theme_classic()  + geom_vline(xintercept = assign_data$date[75], show.legend = FALSE) + annotate("text", x=assign_data$date[70], y=86, label="Start of Pandemic ", angle=90) + geom_line() +  scale_x_discrete(breaks = c("2020-01-01",
                              "2020-02-01",
                              "2020-03-01",
                              "2020-04-01",
                              "2020-05-01",
                              "2020-06-01",
                              "2020-07-01",
                              "2020-08-01")) + ggtitle("Interest in Food Types Over Time")

hq1_line_graph

# Look at plot at pre-treatment, to gauge trend lines and similarities or differences: 

hq1_line_graph + coord_cartesian(xlim= c(0,76))
#we can see before the start of the pandemic sourdough looked similar to cereal but did not look anything like soup or sandwich so considering droping thos



```
Based on the graph we created, when the pandemic started there were temporary effects on the popularity of sourdough bread. From the graph we can see that after the start of the pandemic the hits for sourdough increased to a certain point then started to decline back to pre pandemic levels. For cereal and sandwich respectively, their popularity levels remained relatively the same before and after the pandemic started. Soup was experiencing a decline in popularity before the pandemic began and then experienced a small increase in popularity when the pandemic began but soon after continued its descent in popularity. As sourdough is our variable of interest, it was interesting to see that the levels in popularity increased for a few months after the pandemic began however it appears that as people got used to living in the pandemic, they returned to their pre pandemic habits. 

Sourdough bread will be considered our Treated group and thus when looking at the other foods as controls, some are more appropriate than others. When the graph was zoomed in to look at the popularity trends before the pandemic began it can be noted that the trend of sourdough bread looks most like that of cereal. They both experience low peaks and dips in their popularity trend line and have very similar locations at where these peaks and dips occur. Sourdough's trend line and sandwich's trend line do not look as similar, as sandwich popularity appears to have much more changes and sharper dips and peaks than that of sourdough's. The trend line for soup is the most distinct from that of sourdoughs as it has large variation in popularity over time and it has a clear downard trend in popularity. Based on the pre-pandemic trend lines, it appears that cereal would serve as the best control group for sourdough as the two trend lines look very similar and thus without any impactful events (such as a pandemic) it could be predicted that their trend lines would look similar if extrapolated. 


# Question 1b
```{r Question 1b}

#Create Treated Variable for sourdough https://www.princeton.edu/~otorres/DID101R.pdf

regdata <- assign_data %>% mutate(treated = ifelse(keyword == "sourdough", 1, 0)) 

# Create a dummy variable to indicate the time when the treatment started. 

regdata <- regdata %>% mutate(pandemic_begun = ifelse(date >= "2020-03-15", 1, 0))

#Create interaction variable between treated and beginning of pandemic

regdata <- regdata %>% mutate(interaction = treated*pandemic_begun)

# Statistical tests: are the two groups' trends different?
pretrend_test <- lm(hits ~ treated + pandemic_begun + interaction, data=regdata)
msummary(list(pretrend_test),
         vcov=c(rep("robust",1)),
         stars=c('*' = .1, '**' = .05, '***' = .01)) # Interpret each coefficient here

#Dropping Control Groups

#dropping soup and sandwich from data 
new_regdata<- regdata %>% filter(keyword %in% c("sourdough","cereal"))

unique(new_regdata$keyword)

new_pretrend_test <- lm(hits ~ treated + pandemic_begun + interaction, data=new_regdata)
msummary(list(pretrend_test,new_pretrend_test),
         vcov=c(rep("robust",1)),
         stars=c('*' = .1, '**' = .05, '***' = .01)) # Interpret each coefficient here

#i think this is how you do it but im not sure


```

# Question 1c
```{r Question 1c}

#create a realtive_month variable by shifting the "date" variable back 15 days and then taking the month of the resulting date



#create an "After" variable equal to 1/0 or True/False if the date is March 15 or afterwards 





#illustrate how the values of "relative_month" line up with "date" and subtracte a number from "relative_month" so the last period just before treatment (Feb 16 - Mar 14)  is 0. (Also, change the Jan 1 -14 month so it's one less than the Jan 15- Feb 14 month) 



```

# Question 1d
```{r Question 1d}

```




